 

### 2023-03-11 appendix.Rmd
Revisions throughout and final version for External Reader

col_CH <- "coral"
col_CHD <- "coral1"
col_CHR <- "coral3"
col_CSB <- "steelblue2"
col_CSB1 <- "steelblue1"
col_CSB2 <- "steelblue3"
col_other <-  "grey20"

>>>>

col_CH <- "coral"
col_CHD <- "coral1"
col_CHR <- "deeppink3"
col_CSB <- "steelblue2"
col_CSB1 <- "steelblue1"
col_CSB2 <- "steelblue3"
col_other <-  "grey20"

### 2023-03-11 appendix_2.1.Rmd
Revisions throughout and final version for External Reader

The health facility data used in our study was obtained from Madagascar's *Gestion du Système d'Information Sanitaire* (GESIS), a Microsoft Access electronic health management information systems database.

While primary, district-level, and regional referral facilities in Madagascar all report monthly case counts of mental disorders in GESIS, case definitions vary across facilities of different levels. Disaggregated data by specific condition (e.g., depression, schizophrenia, etc.) are only reported among regional referral hospitals and university hospitals; for all other facilities, cases are reported in aggregate (e.g., "Neuro-psychic diseases" for district hospitals).

Disaggregated data is available by age; among adults, however, disaggregated data for adults ages 25 to 59 and adults 60 years and older has only been reported by primary health facilities since 2015. For all other years and all other facilities, these age categories are not separately reported. Similarly, data by sex is only reported for primary health facilities since 2015. Referrals are reported by primary health facilities and district hospitals.

A summary of the GESIS data availability is provided in Table S2.1.

@@

@@

We then conducted exploratory data analysis to gauge data availability by condition, year, and facility level, which are reported in Tables S2.3-5. These tables show yearly cumulative cases of mental and neurological disorders, by facility level.

@@

Among district hospitals, reporting of mental disorders has been consistent between 2010 and 2020.

Reporting of mental disorders is most detailed among regional referral hospitals (CHRRs) and university hospitals (CHUs), with case counts reported for `r  prettyNum(n_chu, big.mark = ",", scientific = FALSE)` unique conditions.

@@

\addcontentsline{lot}{table}{Table S2.3: Yearly cumulative cases of mental and neurological disorders reported at primary facilities.}

@@

reporting_table(csb_data, "Table S2.3: Yearly cumulative cases of mental and neurological disorders reported at primary facilities")

@@

\addcontentsline{lot}{table}{Table S2.4: Yearly cumulative cases of mental and neurological disorders reported at district hospitals.}

@@

reporting_table(chd_data, "Table S2.4: Yearly cumulative cases of mental and neurological disorders reported at district hospitals")

@@

\addcontentsline{lot}{table}{Table S2.5: Yearly cumulative cases of mental and neurological disorders reported at regional hospitals.}

@@

reporting_table(chu_data, "Table S2.5: Yearly cumulative cases of mental and neurological disorders reported at regional hospitals")

@@

To address heterogeneity in reporting across primary, district-level, and regional-level facilities, our analysis constructed a single, summary outcome measure: any mental or neurological disorders reported in a given month by a given facility, irrespective of age and sex. This outcome measure was constructed by grouping our count data by facility and month-year, and then summing the incidence of mental and neurological disorders reported in GESIS within groups, preventing double counting and the inclusion of referrals in the summary measure.

@@

data_mh %>%
  filter(cleaning_status == "count_raw") %>%
  ggplot(aes(x = as.Date(date), y = count)) +
  geom_point(alpha = 0.05, aes(color = clinic_type_clean)) +
    geom_smooth(colour = col_black, fill = col_grey_med) +
  scale_y_log10() +
    theme_classic() +
  theme(strip.background = element_rect(colour="white",
                                        fill="white"),
        text = element_text(family = "Source Sans Pro")) +
  theme(legend.position = "none") +
  labs(x = "Date", y = "Monthly count of cases") +
  facet_wrap(~clinic_type_clean, nrow = 3) +
  scale_color_manual(values = c(col_CHD, col_CHR, col_CSB))

@@

Visual inspect of our data showed potential outliers. which may have arisen due to data entry mistakes, synchronization errors, and other sources. To address outliers, monthly observations of the outcome were arranged separately for each facility into a unique time series. The nonparametric Friedman’s super smoother regression estimator [@friedmanVariableSpanScatterplot1984] was applied to each time series to identify outlier observations and replace them through linear interpolation [@hyndmanAutomaticTimeSeries2008]. By considering outliers on a facility-by-facility basis, the data cleaning approach preserves extreme values that may reflect true variances in incidence (for example, incidences reported by a large, regional reference hospital) while addressing potential errors in data entry. Interpolated values were coerced to the nearest integer.

>>>>

The health facility data used in our study were obtained from Madagascar's *Gestion du Système d'Information Sanitaire* (GESIS), a Microsoft Access electronic health management information systems database.

While primary, district, and regional referral facilities in Madagascar all reported monthly visits related to mental disorders in GESIS, case definitions varied across facilities of different levels. Disaggregated data by specific condition (e.g., depression, schizophrenia, etc.) were only reported among regional referral hospitals and university hospitals; for all other facilities, visit by type of mental heath complaint were reported in aggregate (e.g., "Neuro-psychic diseases" for district hospitals).

Disaggregated data was available by age; among adults, however, disaggregated data for adults ages 25 to 59 and adults 60 years and older was only reported by primary health facilities after 2015. For all other years and all other facilities, these age categories were not separately reported. Similarly, data by sex was only reported for primary health facilities after 2015. Referrals were reported by primary health facilities and district hospitals.

A summary of GESIS data availability is provided in Table S2.1.

@@

\vspace{6 pt}

@@

We then conducted exploratory data analysis to gauge data availability by condition, year, and facility level, which are reported in Tables S2.3-5. These tables show yearly cumulative visits for mental and neurological disorders, by facility level.

@@

Among district hospitals, reporting of mental disorders was consistent between 2010 and 2020.

Reporting of mental disorders was most detailed among regional referral hospitals (CHRRs) and university hospitals (CHUs), with visit counts reported for `r  prettyNum(n_chu, big.mark = ",", scientific = FALSE)` unique conditions.

@@

\addcontentsline{lot}{table}{Table S2.3: Yearly cumulative visits for mental and neurological disorders reported at primary facilities.}

@@

reporting_table(csb_data, "Table S2.3: Yearly cumulative visits for mental and neurological disorders reported at primary facilities")

@@

\addcontentsline{lot}{table}{Table S2.4:  Yearly cumulative visits for mental and neurological disorders reported at district hospitals.}

@@

reporting_table(chd_data, "Table S2.4: Yearly cumulative visits for mental and neurological disorders reported at district hospitals")

@@

\addcontentsline{lot}{table}{Table S2.5: Yearly cumulative visits for mental and neurological disorders reported at regional hospitals.}

@@

reporting_table(chu_data, "Table S2.5: Yearly cumulative visits for mental and neurological disorders reported at regional hospitals")

@@

To address heterogeneity in reporting across primary, district, and regional facilities, our analysis constructed a single, summary outcome measure: any mental or neurological disorders reported in a given month by a given facility, irrespective of age and sex. This outcome measure was constructed by grouping our count data by facility and month-year, and then summing the incidence of mental and neurological disorders reported in GESIS within groups, preventing double counting and the inclusion of referrals in the summary measure.

@@

data_mh %>%
  filter(cleaning_status == "count_raw") %>%
  ggplot(aes(x = as.Date(date), y = count)) +
  geom_point(alpha = 0.05, aes(color = clinic_type_clean)) +
    geom_smooth(colour = col_black, fill = col_grey_med) +
  scale_y_log10() +
    theme_classic() +
  theme(strip.background = element_rect(colour="white",
                                        fill="white"),
        text = element_text(family = "Source Sans Pro")) +
  theme(legend.position = "none") +
  labs(x = "Date", y = "Monthly count of visits") +
  facet_wrap(~clinic_type_clean, nrow = 3) +
  scale_color_manual(values = c(col_CHD, col_CHR, col_CSB))

@@

Visual inspect of our data showed potential outliers, which may have arisen due to data entry mistakes, synchronization errors, and other sources. To address outliers, monthly observations of the outcome were arranged separately for each facility into a unique time series. The nonparametric Friedman’s super smoother regression estimator [@friedmanVariableSpanScatterplot1984] was applied to each time series to identify outlier observations and replace them through linear interpolation [@hyndmanAutomaticTimeSeries2008]. By considering outliers on a facility-by-facility basis, the data cleaning approach preserved extreme values that may habe reflected true variances in visit counts (for example, visits reported by a large, regional reference hospital) while addressing potential errors in data entry. Interpolated values were coerced to the nearest integer.

### 2023-03-11 appendix_2.2.Rmd
Revisions throughout and final version for External Reader

n_facilities_name <- data_mh %>% dplyr::select(clinic_name) %>% unique %>% nrow()
n_facilities <- data_mh %>% unique %>% nrow()

@@

In Madagascar, public health facilities can be broadly classified under four types:

@@

The sample of health facilities in our study was obtained from health facility data from Madagascar's *Gestion du Système d'Information Sanitaire* (GESIS). While the format of outputs from GESIS vary across years and facility type, there are four variables which uniquely identify facilities.

* **Region**: The region of Madagascar that contains the facility's catchment area. All facilities have a value for this variable.
* **District**: The district of Madagascar (2nd largest administrative unit) that contains the facility's catchment area. Only district- and commune-level facilities will have a value for this variable; regional-level hospitals  have an "*NA*" for this variable.
* **Commune**: The commune of Madagascar (3rd largest unit) that contains the facility's catchment area. District- and regional-level hospitals have an "*NA*" for this variable.
* **FS**: The facility type and name. The facility type will typically be identified with an acronym (e.g., CSB2, CHD1, etc.), with the facility name following thereafter. The facility name will occasionally be the same as the commune, district, or region that corresponds with the facility's catchment area (e.g., CHRR Alaotra Mangoro). Some facilities are named for the village (*fokontany*) that they are located in. Other facilities are named for the groups that operate them (e.g., facilities named "EKAR" are operated by the *Eglizy Katôlika Apôstôlika Rômanina*, or the Roman Catholic Church).

@@

Using facility name, type, and catchment area to uniquely identify facilities in our sample, we identified `r  prettyNum(n_facilities, big.mark = ",", scientific = FALSE)` unique facilities. In order to classify these facilities by type, we extracted the relevant data from the "FS' variable; in doing so, we observed slight orthographic variations in the classification of facilities types (e.g., "CSB 1" vs. "CSB1") that required standardization.

We further standardized region, district, and commune names in our sample against reference shapefiles to address variations in translation (e.g., "Atsimo" and "Sud" being the Malagasy and French words for "South"), abbreviation (e.g., "St." as an abbreviation for "Saint"), transposition (e.g., "Bevoay Beretra" transposed as "Beretra Bevoay"), accent marks (e.g., "Tovòna" vs. "Tovona"), spacing (e.g., "AmbalapaisoII" vs. "Ambalapaiso II"), separation vs. combination of names (e.g., "Tanambaovatrakaka" vs. "Tanambao Vahatrakaka"), hyphenation (e.g, "Ankiabe Salohy" and "Ankiabe-Salohy"), and spelling (e.g., "Mizilo Gare" vs.	"Mizilo Gara").

@@

After standardizing facility types and catchment areas, the sample consisted of `r  prettyNum(n_facilities_combined, big.mark = ",", scientific = FALSE)` unique facilities. This list of facilities was then manually inspected for any remaining duplicates. From reviewing this list, we observed that in the Diana region, the regional hospitals named "Hopitaly MANARA-PENITRA" and "Hopitaly Manara-Penitra Antsiranana" are duplicates, but appear separately because the later name contains the city the hospital is located in, Antsiranana. We standardized the names of those facilities, which resolved `r  prettyNum(n_dif, big.mark = ",", scientific = FALSE)` duplicate(s).

The final sample of `r  prettyNum(n_facilities_new, big.mark = ",", scientific = FALSE)` unique facilities consisted of:

*  `r  prettyNum(n_CSB +n_CSB1 +n_CSB2, big.mark = ",", scientific = FALSE)` CSB facilities, including `r  prettyNum(n_CSB1, big.mark = ",", scientific = FALSE)` CSB1s, `r  prettyNum(n_CSB2, big.mark = ",", scientific = FALSE)` CSB2s, and `r  prettyNum(n_CSB, big.mark = ",", scientific = FALSE)` of unknown level;
*   `r  prettyNum(n_FSP, big.mark = ",", scientific = FALSE)`  FSP facilities;
*  `r  prettyNum(n_CHD + n_CHD1 + n_CHD2, big.mark = ",", scientific = FALSE)`  district hospitals, including `r  prettyNum(n_CHD1, big.mark = ",", scientific = FALSE)` CHD1 facilities, `r  prettyNum(n_CHD2, big.mark = ",", scientific = FALSE)` CHD2 facilities, and `r  prettyNum(n_CHD, big.mark = ",", scientific = FALSE)` facilities of unknown level, and
*  `r  prettyNum(n_CHRR +n_CHU +n_CHRRCHU, big.mark = ",", scientific = FALSE)` regional hospitals, including `r  prettyNum(n_CHRR, big.mark = ",", scientific = FALSE)` CHRRs, `r  prettyNum(n_CHU, big.mark = ",", scientific = FALSE)` CHUs, and `r  prettyNum(n_CHRRCHU, big.mark = ",", scientific = FALSE)` of unknown type.

@@

**Step 1: Matching to validated spatial inventory of public facilities in Sub-Saharan Africa (Maina et al.)**
``` {r step_1-loading_data, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE}

@@

To maximize pair accuracy, we separately conducting matching among primary health facilities (i.e., FSPs and CSBs) and among hospitals (i.e., CHDs, CHRRs, and CHUs). For primary health facilities, the full, standardized facility names included the facility's region, district, and commune; for hospitals, the full name included the region and district only. To apply a conservative matching approach, we restricted matches to those with a Levenshtein distance no greater than 2.

@@

Using this approach, `r format(as.numeric(n_matched_facilities_rhino), big.mark = ",")` facilities were matched to coordinates in the RHINoVision spatial inventory, of which `r format(as.numeric(n_adist_rhino_0), big.mark = ",")` facilities were matched based on identical string matches,  `r format(as.numeric(n_adist_rhino_1), big.mark = ",")` facilities were matched based on string matches with a Levenshtein distance of 1, `r format(as.numeric(n_adist_rhino_2), big.mark = ",")` facilities were matched based on string matches with a Levenshtein distance of 2, and `r format(as.numeric(n_adist_rhino_3), big.mark = ",")` facilities were matched based on string matches with a Levenshtein distance of 3. (`r format(as.numeric(n_adist_rhino_0+n_adist_rhino_1+n_adist_rhino_2+n_adist_rhino_3 - n_matched_facilities_rhino), big.mark = ",")` matches were duplicates and were removed.)

@@

@@

Most of the coordinates in this expanded inventory were duplicates (i.e., the same facility will appear in different sources). Importantly, coordinates for the same facility could be slightly different across sources, due to variation in precision and geolocation.

@@

#

@@

                                    # 
                                    # 
                                    # ifelse(length == 0,
                                    #        "polygon", "other")))) %>%

  # Let's apply our thresholds from the PARENT file
  mutate(retain = ifelse(length == 0 & area_km == 0, 
                         "Within threshold",
                         ifelse(length <= length_lim & area_km == 0, 
                                "Within threshold",
                                ifelse(length == 0 & area_km <= area_lim,
                                       "Within threshold", 
                                       "Above threshold")))) %>%
  ungroup()

@@

We applied an arbitrary restriction to line and polygon groups, such that the length or area, respectively, would not exceed a particular threshold. The threshold for length (`r format(as.numeric(length_lim), big.mark = ",")` km.) was selected to be approximately half the distance of one side of a spatial climate resolution tile (~31 km for ERA-5 data). The threshold for area (`r format(as.numeric(area_lim), big.mark = ",")` square km.) was assigned to be the length threshold, squared. Groups that exceeded their respective thresholds were discarded.

@@

Among polygon groups, the mean area was `r format(round(n_mean_area_polygons,2), big.mark = ",")` km², with a maximum area of `r format(round(n_max_area_polygons), big.mark = ",")` km². At the threshold of `r format(as.numeric(area_lim), big.mark = ",")` square km., all polygon groups fell within the cut-off. Among line groups, the mean distance between coordinates was `r format(round(n_mean_length,2), big.mark = ",")` km, with a maximum distance of `r format(round(n_max_length), big.mark = ",")` km. At the threshold of `r format(as.numeric(length_lim), big.mark = ",")` km,  `r format(round(n_length_retained), big.mark = ",")` line groups fell within the cut-off (`r round(as.numeric(n_length_retained)/n_line * 100,0)`% of line groups).

@@

While it would be ideal to obtain precise locations for each facility in our sample, for the purposes of our analysis, an approximate location would be sufficient, given the spatial resolution of the climate data we used. The spatial resolution of ERA-5 and ESA-CCI data is approximately 30 km., meaning that all coordinates within a spatial tile of 30 km² will have the same time series for a given variable. Furthermore, neighboring tiles are unlikely to have extraordinarily different time series, underscoring that close approximate locations would likely result in minimal impact to the analysis versus exact coordinates.

For commune-level facilities, knowing the catchment area may therefore be just as informative as knowing the facility's precise location if the catchment area was sufficiently small. Determining a cut-off was, again, arbitrary; to adopt a conservative threshold, we considered a catchment area small if its area was  no greater than `r format(as.numeric(area_lim), big.mark = ",")` square km, which corresponded to approximately `r round(as.numeric(area_lim)/(30*30) * 100,0)`% of the area of a climate data spatial resolution tile (Figure \ref{figure_communes_fokontany}).

@@

To exploit all available information for geolocation, we first attempted to match commune-level facilities to known *fokontany* based on the facility name and catchment area. We used approximate string matching to link facilities named for their locations in our sample to known villages; if the village area was smaller than our threshold, the centroid of the village was used as a proxy for the facility location. We then applied the same approach to unmatched commune-level facilities to known communes.

@@

To verify the accuracy of the geolocating process, we randomly selected `r round(as.numeric(sample_fraction * 100,0))`% of our sample of geolocated facilities (n = `r format(as.numeric(n_facilities_sample), big.mark = ",")`) and independently identified coordinates for each facility. We then compared these coordinates to those identified from our geolocating process. We then calculated the physical distance between the independently identified coordinates and the coordinates identified from our geolocating process for each facility.

Table S2.10 provides the list of facilities in the validation sample.

@@

* *Andrefana*: west (Malagasy)
* *Atsimo*: south (Malagasy)
* *Atsinanana*: east (Malagasy)
* *Avaratra*: north (Malagasy)
* *Est*: west (French)
* *Haute*: high (French)
* *Nord*: north (French)
* *Ouest*: west (French)
* *Sud*: south (French)
* *Vaovao*: new (Malagasy)

>>>>

n_facilities_name <- data_mh %>% dplyr::select(clinic_name) %>% unique %>% nrow()
n_facilities_overall <- data_mh %>% unique %>% nrow()

@@

In Madagascar, public health facilities can broadly be classified under four types:

@@

The sample of health facilities in our study was obtained from health facility data from Madagascar's *Gestion du Système d'Information Sanitaire* (GESIS). While the format of outputs from GESIS varied across years and facility type, there were four variables which uniquely identified facilities.

* **Region**: The region of Madagascar that contained the facility's catchment area. All facilities had a value for this variable.
* **District**: The district of Madagascar (2nd largest administrative unit) that contained the facility's catchment area. Only district- and commune-level facilities had a value for this variable.
* **Commune**: The commune of Madagascar (3rd largest unit) that contained the facility's catchment area. Only basic health facility had a value for this variable.
* **FS**: The facility type and name. The facility type would typically be identified with an acronym (e.g., CSB2, CHD1, etc.), with the facility name following thereafter. The facility name would occasionally be the same as the commune, district, or region that corresponded with the facility's catchment area (e.g., CHRR Alaotra Mangoro). Some facilities were named for the village (*fokontany*) that they were located in. Other facilities were named for the groups that operated them (e.g., facilities named "EKAR" are operated by the *Eglizy Katôlika Apôstôlika Rômanina*, or the Roman Catholic Church).

@@

Using facility name, type, and catchment area to uniquely identify facilities in our sample, we identified `r  prettyNum(n_facilities_overall, big.mark = ",", scientific = FALSE)` unique facilities. In order to classify these facilities by type, we extracted the relevant data from the "FS' variable; in doing so, we observed slight orthographic variations in the classification of facilities types (e.g., "CSB 1" vs. "CSB1") that required standardization.

We standardized region, district, and commune names in our sample against reference shapefiles to address variations in translation (e.g., "Atsimo" and "Sud" being the Malagasy and French words for "South"), abbreviation (e.g., "St." as an abbreviation for "Saint"), transposition (e.g., "Bevoay Beretra" transposed as "Beretra Bevoay"), accent marks (e.g., "Tovòna" vs. "Tovona"), spacing (e.g., "AmbalapaisoII" vs. "Ambalapaiso II"), separation vs. combination of names (e.g., "Tanambaovatrakaka" vs. "Tanambao Vahatrakaka"), hyphenation (e.g, "Ankiabe Salohy" and "Ankiabe-Salohy"), and spelling (e.g., "Mizilo Gare" vs.	"Mizilo Gara").

@@

After standardizing facility types and catchment areas, the sample consisted of `r  prettyNum(n_facilities_combined, big.mark = ",", scientific = FALSE)` unique facilities. This list of facilities was then manually inspected for any remaining duplicates. From reviewing this list, we observed that in the Diana region, the regional hospitals named "Hopitaly MANARA-PENITRA" and "Hopitaly Manara-Penitra Antsiranana" were duplicates, but appeared separately because the latter name contained the city the hospital is located in, Antsiranana. We standardized the names of those facilities, which resolved `r  prettyNum(n_dif, big.mark = ",", scientific = FALSE)` `r ifelse(n_dif > 1, "duplicates","duplicate")`).

The final sample of `r  prettyNum(n_facilities_new, big.mark = ",", scientific = FALSE)` unique facilities consisted of:  `r  prettyNum(n_CSB +n_CSB1 +n_CSB2, big.mark = ",", scientific = FALSE)` CSB facilities, including `r  prettyNum(n_CSB1, big.mark = ",", scientific = FALSE)` CSB1s, `r  prettyNum(n_CSB2, big.mark = ",", scientific = FALSE)` CSB2s, and `r  prettyNum(n_CSB, big.mark = ",", scientific = FALSE)` of unknown level; `r  prettyNum(n_FSP, big.mark = ",", scientific = FALSE)`  FSP facilities; `r  prettyNum(n_CHD + n_CHD1 + n_CHD2, big.mark = ",", scientific = FALSE)`  district hospitals, including `r  prettyNum(n_CHD1, big.mark = ",", scientific = FALSE)` CHD1 facilities, `r  prettyNum(n_CHD2, big.mark = ",", scientific = FALSE)` CHD2 facilities, and `r  prettyNum(n_CHD, big.mark = ",", scientific = FALSE)` facilities of unknown level, and `r  prettyNum(n_CHRR +n_CHU +n_CHRRCHU, big.mark = ",", scientific = FALSE)` regional hospitals, including `r  prettyNum(n_CHRR, big.mark = ",", scientific = FALSE)` CHRRs, `r  prettyNum(n_CHU, big.mark = ",", scientific = FALSE)` CHUs, and `r  prettyNum(n_CHRRCHU, big.mark = ",", scientific = FALSE)` of unknown type.

@@

**Step 1: Matching to validated spatial inventory of public facilities in Sub-Saharan Africa**
``` {r step_1-loading_data, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE}

@@

To maximize pair accuracy, we separately conducted matching among primary health facilities (i.e., FSPs and CSBs) and among hospitals (i.e., CHDs, CHRRs, and CHUs). For primary health facilities, the full, standardized facility names included the facility's region, district, and commune; for hospitals, the full name included the region and district only. To apply a conservative matching approach, we restricted matches to those with a Levenshtein distance no greater than 2.

@@

library(english)

Using this approach, `r format(as.numeric(n_matched_facilities_rhino), big.mark = ",")` facilities were matched to coordinates in the RHINoVision spatial inventory, of which `r format(as.numeric(n_adist_rhino_0), big.mark = ",")` facilities were matched based on identical string matches,  `r format(as.numeric(n_adist_rhino_1), big.mark = ",")` facilities were matched based on string matches with a Levenshtein distance of 1, `r format(as.numeric(n_adist_rhino_2), big.mark = ",")` facilities were matched based on string matches with a Levenshtein distance of 2, and `r format(as.numeric(n_adist_rhino_3), big.mark = ",")` facilities were matched based on string matches with a Levenshtein distance of 3. (`r str_to_sentence(english(as.numeric(n_adist_rhino_0+n_adist_rhino_1+n_adist_rhino_2+n_adist_rhino_3 - n_matched_facilities_rhino)))` matches were duplicates and were removed.)

@@

facilities_gps_any_test <- facilities_gps_any

@@

Most of the coordinates in this expanded inventory were duplicates, as the same facility would appear in different sources. Importantly, coordinates for the same facility could be slightly different across sources, due to variation in precision and geolocation.

@@

sf_use_s2(TRUE)

@@

  # Let's apply our thresholds from the PARENT file
  mutate(retain = 
           ifelse(shape_type == "point" |
                    (shape_type == "line" & length <= length_lim) |
                    (shape_type == "polygon" & area_km <= area_lim),
                                       "Within threshold", 
                                       "Above threshold")) %>%
  ungroup()

@@

We applied an arbitrary restriction to line and polygon groups, such that the length or area, respectively, would not exceed a particular threshold. The threshold for length (`r format(as.numeric(length_lim), big.mark = ",")` km.) was selected to be approximately half the distance of one side of a spatial climate resolution tile (~30 km for ERA-5 data). The threshold for area (`r format(as.numeric(area_lim), big.mark = ",")` km²) was assigned to be the length threshold, squared. Groups that exceeded their respective thresholds were discarded.

@@

Among polygon groups, the mean area was `r format(round(n_mean_area_polygons,2), big.mark = ",")` km², with a maximum area of `r format(round(n_max_area_polygons), big.mark = ",")` km². At the threshold of `r format(as.numeric(area_lim), big.mark = ",")` km², all polygon groups fell within the cut-off. Among line groups, the mean distance between coordinates was `r format(round(n_mean_length,2), big.mark = ",")` km, with a maximum distance of `r format(round(n_max_length), big.mark = ",")` km. At the threshold of `r format(as.numeric(length_lim), big.mark = ",")` km,  `r format(round(n_length_retained), big.mark = ",")` line groups fell within the cut-off (`r round(as.numeric(n_length_retained)/n_line * 100,0)`% of line groups).

@@

While it would be ideal to obtain precise locations for each facility in our sample, for the purposes of our analysis, an approximate location would be sufficient, given the spatial resolution of the climate data we used. The spatial resolution of ERA-5 and ESA-CCI data is approximately 30 km., meaning that all coordinates within a spatial tile of 900 km² will have the same time series for a given variable. Furthermore, neighboring tiles are unlikely to have extraordinarily different time series, underscoring that close approximate locations would likely result in minimal impact to the analysis versus exact coordinates.

For commune-level facilities, knowing the catchment area may therefore be just as informative as knowing the facility's precise location if the catchment area was sufficiently small. Determining a cut-off was, again, arbitrary; to adopt a conservative threshold, we considered a catchment area small if its area was  no greater than `r format(as.numeric(area_lim), big.mark = ",")` km², which corresponded to approximately `r round(as.numeric(area_lim)/(30*30) * 100,0)`% of the area of a climate data spatial resolution tile (Figure \ref{figure_communes_fokontany}).

@@

To exploit all available information for geolocation, we first used approximate string matching to link facilities named for their locations in our sample to known *fokontany*, based on the facility name and catchment area; if the village area was smaller than our threshold, the centroid of the village was used as a proxy for the facility location. We then applied the same approach to unmatched commune-level facilities to known communes.

@@

To verify the accuracy of the geolocating process, we randomly selected `r round(as.numeric(sample_fraction * 100,0))`% of our sample of geolocated facilities (n = `r format(as.numeric(n_facilities_sample), big.mark = ",")`) and independently identified coordinates for each facility. We then compared these coordinates to those identified from our geolocating process. We then calculated the physical distance between the independently identified coordinates and the coordinates identified from our geolocating process for each facility. Table S2.10 provides the list of facilities in the validation sample.

@@

* *Andrefana*: west (Malagasy)
* *Atsimo*: south (Malagasy)
* *Atsinanana*: east (Malagasy)
* *Avaratra*: north (Malagasy)
* *Est*: east (French)
* *Haute*: high (French)
* *Nord*: north (French)
* *Ouest*: west (French)
* *Sud*: south (French)
* *Vaovao*: new (Malagasy)

### 2023-03-11 conclusion.Rmd
Revisions throughout and final version for External Reader

Overview of Paper 1

Overview of Paper 2

Overview of Paper 3

Policy implications and future research

The magnitude of this burden has galvanized a global movement and a call to action for greater investment and prioritization for mental health.[@patelRenewedAgendaGlobal2011] This movement has emphasized the importance of investing in mental health as a means of promoting sustainable development, human rights, and social inclusion.[@patelLancetCommissionGlobal2018]

Burden may be larger than traditionally conceptualized

Papers also show that mental health is linked to broader contextual challenges (climate change, COVID-19) – resilience/vulnerability, exacerbating existing disparities, but also could be critical to overcoming them…

In the absence of population mental health metrics, innovative uses of health data may be helpful in conducting research, but come with limitations

>>>>

In this dissertation, I investigated different quantitative facets of the global burden of mental disorders, a burden that is widespread,[@charlsonExcessMortalityMental2015; @rehmGlobalBurdenDisease2019] growing,[@patelLancetCommissionGlobal2018] and largely unmet.[@bassUnitedNationsGeneral2012a; @demyttenaerePrevalenceSeverityUnmet2004]

In Chapter 1, we examined how the global burden is defined, and we proposed an estimation approach to more comprehensively reflect morbidity and loss of life due to suicide and premature mortality attributable to mental disorders. After applying this composite estimation approach to the most recently available estimates from the Global Burden of Disease study, we found that approximately one in 6 disability-adjusted life years (approximately 16%) were due to mental disorders in 2019. We then applied a value of a statistical life year approach to estimate the associated global economic burden. Our findings suggest staggering loses: after accounting for premature mortality and other sources of morbidity associated with mental disorders, global losses were estimated to approximate 4.7 trillion USD in 2019. The magnitude of these findings underscores the global scale of impact that mental disorders have and highlights the urgency for research and action to support mental health worldwide.

In Chapter 2, we examined how this global burden might be shifting with changes in climate by analyzing health system data in Madagascar, a country with little to no mental health surveillance data and high vulnerability to climate change. We obtained data on monthly counts of mental health related visits reported by 3,413 facilities in Madagascar from 2010 to 2020. We then geolocated these facilities and obtained meteorologic data on climate conditions to examine the impact of three climate change exposures—temperature, soil moisture, and duration of cyclones and tropical storms. We found that monthly mental health related visits declined as temperature increased, particularly in the central highlands, where cooler temperatures were more prevalent; conversely, higher grades of soil moisture were associated with greater mental health visits with a lag of approximately three months, an increase found primarily in regions associated with higher flood risk in Madagascar. Our findings illustrate potential heterogeneity in how changing climate conditions may impact mental health needs, and our work demonstrates how routine health system data may be leveraged to address data gaps in health surveillance, particularly in resource-constrained settings.

In Chapter 3, I examined how search data for health symptoms could, in the absence of high frequency surveillance data, provide insight into changes in population mental health during the COVID-19 pandemic. I obtained Google search data from U.S. counties, states, and international trends for over 400 symptoms and used these data to study the immediate and sustained impacts of announcements of the development of safe and effective COVID-19 vaccines in November 2020. I found that searches for anxiety, depression, and major depressive disorder significantly fell after these announcements, and that this decline continued throughout the subsequent weeks. Furthermore, despite the subsequent (and ongoing) politicization of COVID-19 vaccines in the United States, I observed consistency in search density declines for mental health symptoms across red and blue states/counties. The findings speak to the importance of timely and transparent communication in public health, as well as to the potential use of internet search data to investigate changes in population mental health that may be difficult to detect without high frequency data.

These papers add to a growing understanding of the tremendous scale of mental health needs worldwide. The magnitude of this burden has galvanized a global movement and a call to action for greater investment and prioritization for mental health.[@patelRenewedAgendaGlobal2011] This movement has emphasized the importance of investing in mental health as a means of promoting sustainable development, human rights, and social inclusion.[@patelLancetCommissionGlobal2018]

To respond to this burden, health systems will need to be mindful of factors beyond the traditional extent of mental health services and consider mental health within a background of broader contextual challenges. As seen in Chapters 2 and 3, changes in climate and the course of the COVID-19 pandemic highlight the importance of considering the connections between mental health and other issues in public health. Insofar as these challenges may exacerbate one another and strengthen vulnerabilities and inequities in health, responding to them in concert—and addressing the mental health needs of populations facing new and evolving stressors—may be key in overcoming them and fostering resiliency. This will require global, multisectoral action across countries at all economic levels, coupled with a substantive increase in financial resources dedicated to mental health.

Moreover, health systems and global health research will need to develop new means of dealing with the current limitations in the measurement of population mental health. As discussed in Chapter 1, traditional approaches to estimating the global burden of mental disorders may not capture its true scale, contributing to a lack of perceived importance of mental health to human well-being. This perception remains unchallenged, in part, due to the lack of detailed epidemiological data on the prevalence and distribution of mental disorders, a challenge that further constrains mental health research. As demonstrated by Chapters 2 and 3, innovative uses of health system and internet search data may help in overcoming this constraint but come with limitations of their own.

While this dissertation has underscored that the burden of mental disorders is complex in nature and vast in scope, the challenges this burden poses are not insurmountable. By measuring and understanding the mental health needs of populations, and by addressing these needs within a broader context of societal challenges, health systems may more effectively respond to this challenge, bringing the world closer to parity between mental health and physical health. Through a global effort to address this essential facet of human wellbeing, millions of individuals may be helped to achieve their full potential, strengthen resilience, and obtain access to the care they need to flourish.
